---
title: Take A Sad Plot And Make It Better 2.0 - Graph 3
author: Ajinkya Shinde
date: '2018-06-16'
slug: sad-plot-better-3
categories: []
tags:
  - ggplot2
  - dplyr
  - tidyverse
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE , error= FALSE, message=FALSE,echo=FALSE}
library(here) # to set file path if working from local file
library(tidyverse) # 
library(tidyr)
library(readr)
library(ggthemes)
library(extrafont)
loadfonts(device="win")
```


### 1. Relevant R Code


```{r message=FALSE,warning=FALSE}
moma <- read_csv(here("static", "data", "artworks-cleaned.csv"))

f_moma<-moma %>%
  filter(artist_gender == "Female")  

 
f_moma2<-count(f_moma,year_created,artist_gender)


ggplot(f_moma2,aes(x=year_created,y=n))+
  geom_col(colour = "#414a4c",fill="#d26868")+
    scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0))+
  labs(x="Year paintings were created",y="Number of 
       Paintings created",
    title="MoMA, Female paintings over the years")+
  geom_smooth(model=lm,se=FALSE,color="#5e020c")+
  theme_tufte(base_family = "Palatino Linotype") +
  theme(plot.title = element_text(hjust = 0.5)) 


```



### 2. Description of the TYPE of graph.

The graph is a bar chart with the geom_smooth line showing the trend line.

### 3. Description of the DATA you used

The data refers to the moma dataset. In this graph, the final data frame is 
f_moma2(final moma). The quantitative variable is the count of number of paintings created.The qualitative variable is the year during which the paintings were created


### 4. Description of the AUDIENCE you are aiming for 

The audience over here is someone who wants to see the paintings done by women over the years 

### 5. Representation Description:

This graph shows the count of number of paintings done by women across the years.

### 6. How to read it & What to look for

For year-gap on x-axis, read the number of paintings created during that 
period. 


**Major Highlights**
It seems that Women actually painted more paintings during the year 1960-2000. Also,the highest number of paintings(15) around 1985-1990.

### 7. Presentation tips

Using scale_x_continuous and scale_y_continuous, the padding at the starting and ending of the x-axis has been removed. The color scheme used is in relation to the artist-gender(female). The trend line has been added to get the sense of overall increase/decrease using geom_smooth



### 8. Variations and alternatives
This plot can be improved by including the Male gender to see how the Female's performed wrt Male artists.


### 9. How I created it

I formulated the question of what exactly I want: I have to see the diversity in the paintings created as per artist's gender. First, I filtered the data-set on Male and Female. But due to large number of overlapping points , the stacked bar chart wasn't the perfect fit. So, I decided to go with only one gender. 
Due to uncontinuity in the year column, the geom_line wasn't the perfect fit, so decided to go with geom_col

